The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2016-11-28 manuscript/manuscript.Rnw
fix inclusion of housekeeping genes
add derek comments

@@

@@

The
\Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
genes comprising the core placenta transcriptome are shown in
\cref{fig:core_placenta_transcriptome} ordered by their median
transcription level in \ac{FPKM}. We considered a gene to be core to
the placenta transcriptome if a 1:1 ortholog of that gene was
expressed with $\text{FPKM} \ge 10$ in all species studied and was not
a human housekeeping gene according to
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}. Genes which
meet this criteria but are housekeeping genes are depicted in
Supplemental \cref{fig:core_placenta_transcriptome_housekeeping}.
\Cref{fig:core_placenta_transcriptome_shape} shows the number of genes
which would be included in the core placenta set if the \ac{FPKM}
cutoff was set differently; for example, choosing an \ac{FPKM} cutoff
of 1 yields %
\Sexpr{cpt.distribution.shape[`Expression (FPKM)`>=1 & `Min. Exp. in`
  == '14 of 14',max(num_genes)]} %
genes in the core transcriptome. %
In contrast to the shape of the number of genes included in the core
placenta transcriptome, the set of \human housekeeping genes expressed
all placentas is flatter (Supplemental
\cref{fig:core_housekeeping_transcriptome_shape}), and includes more
of the total fraction of genes. However, there is still a large
difference between the set of housekeeping genes which are expressed
in \human and those which are expressed in all placentas.

@@

Supplemental \cref{tab:shape_genes} shows the top 50 genes which are
associated with the shape of placenta attachment, as well as the shape
in which the difference is most significant. The most significant
genes which were associated with diffuse placenta attachment include
\textit{EREG}, a ligand of EGFR, \textit{UNC93A}, a homolog of
\textit{UNC93B1} which is involved in the balancing between Tlr7 and
Tlr9 immune response in
mice~\cite{Brinkmann.ea2007:interactionbetweenERmembrane}, and
\textit{WNT5A}, which is a member of the Wnt ligand family whose
singaling pathway has been implicated in motility in human trophoblast
cells~\cite{Gupta.ea2016:CellSignalingPathwaysInvolved}.

@@

>>>>

  \DEW{The discussion and results seem redundant. Given the title
    shouldn't something about shape be mentioned?}

@@

\DEW{Add gene names}

@@

The
\Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
genes comprising the core placenta transcriptome are shown in
\cref{fig:core_placenta_transcriptome} ordered by their median
transcription level in \ac{FPKM}. We considered a gene to be core to
the placenta transcriptome if a 1:1 ortholog of that gene was
expressed with $\text{FPKM} \ge 10$ in all species studied and was not
a human housekeeping gene according to
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}. Genes which
meet this criteria but are housekeeping genes are depicted in
Supplemental \cref{fig:core_placenta_transcriptome_housekeeping}.
\Cref{fig:core_placenta_transcriptome_shape} shows the number of genes
which would be included in the core placenta set if the \ac{FPKM}
cutoff was set differently; for example, choosing an \ac{FPKM} cutoff
of 1 yields %
\Sexpr{cpt.distribution.shape[`Expression (FPKM)`>=1 & `Min. Exp. in`
  == '14 of 14',max(num_genes)]} %
genes in the core transcriptome. %
In contrast to the shape of the number of genes included in the core
placenta transcriptome, the set of \human housekeeping genes expressed
all placentas is flatter (Supplemental
\cref{fig:core_housekeeping_transcriptome_shape}), and includes more
of the total fraction of genes. However, there is still a large
difference between the set of housekeeping genes which are expressed
in \human and those which are expressed in all placentas.

@@

Supplemental \cref{tab:shape_genes} shows the 50 genes which are most
significantly associated with the shape of placenta attachment, as
well as the shape in which the difference is most significant. The
most significant genes which were associated with diffuse placenta
attachment include \textit{EREG}, a ligand of EGFR, \textit{UNC93A}, a
homolog of \textit{UNC93B1} which is involved in the balancing between
Tlr7 and Tlr9 immune response in
mice~\cite{Brinkmann.ea2007:interactionbetweenERmembrane}, and
\textit{WNT5A}, which is a member of the Wnt ligand family whose
singaling pathway has been implicated in motility in human trophoblast
cells~\cite{Gupta.ea2016:CellSignalingPathwaysInvolved}.

@@

\DEW{Need to cite NWM grants, IGB, etc.}
